---
title: "EC attendances - social care"
author: "NHSEI TD AU"
date: "25/05/2022"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
library(ggplot2)
library(ggthemes)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(fig.width=unit(18,"cm"), fig.height=unit(11,"cm"))

oldw <- getOption("warn")
options(warn = -1)
```


```{css echo=FALSE}
/* Define a margin before h2 element */
h2  {
  margin-top: 2em;
}
/* Define a margin before h3 element */
h3  {
  margin-top: 2em;
}
/* Define a margin before h4 element */
h4  {
  margin-top: 2em;
}
/* Define a margin after every first p elements */
p:first-of-type {
  margin-bottom: 1em;
}
``` 




```{r reference, include=FALSE}

library(tidyverse)
library(kableExtra)

# connect to sandbox
library(DBI)
con <- dbConnect(odbc::odbc(), "NCDR", timeout = 10)

# check which tables are available in the database
#dbListTables(con)

# Querying

# dbGetQuery() will submit query and fetch results
sql_query = "SELECT * FROM [NHSE_Sandbox_EmpowerThePerson].[dbo].[MFonseca_OPArheu02] WHERE Appointment_Type='FUp'"

dataraw <- dbGetQuery(con,
                      sql_query)

### Load reference grouping of chief complaint
data_grouping <- dbGetQuery(con,
                            "SELECT *
                            FROM [NHSE_Reference].[dbo].[tbl_Ref_DataDic_ECDS_Chief_Complaint_Group]")
data_grouping <- data_grouping %>% select(-Last_Refreshed)
data_grouping <- data_grouping %>% mutate(ChiefComplaintCode = as.character(ChiefComplaintCode))


### Load reference complaint description
data_complaint <- dbGetQuery(con,
                             "SELECT [ChiefComplaintCode]
                             ,[ChiefComplaintDescription]
                             FROM [NHSE_Reference].[dbo].[tbl_Ref_DataDic_ECDS_Chief_Complaint]")
data_complaint <- data_complaint %>% mutate(ChiefComplaintCode = as.character(ChiefComplaintCode))


### Load reference attendance source
data_source <- dbGetQuery(con,
                            "SELECT [AttendanceSourceCode],[AttendanceSourceDescription]
                             FROM [NHSE_Reference].[dbo].[tbl_Ref_DataDic_ECDS_Attendance_Source]")
data_source <- data_source %>% mutate(AttendanceSourceCode=as.character(AttendanceSourceCode))
care_source <- c('877171000000103','1077761000000105')
tb_source <- data_source %>% mutate(care = ifelse(AttendanceSourceCode %in% care_source,1,0)) %>% arrange(desc(care))



### Load reference discharge destination
data_discdest <- dbGetQuery(con,
                                    "SELECT [DischargeDestinationCode]
                                    ,[DischargeDestinationDescription]
                             FROM [NHSE_Reference].[dbo].[tbl_Ref_DataDic_ECDS_Discharge_Destination]")
data_discdest <- data_discdest %>% mutate(DischargeDestinationCode=as.character(DischargeDestinationCode))
care_disc <- c('306691003','306694006')
tb_discdest <- data_discdest %>% mutate(care = ifelse(DischargeDestinationCode %in% care_disc,1,0)) %>% arrange(desc(care))

### Load accommodation status
data_accom <- dbGetQuery(con,
                            "SELECT [AccommodationStatusCode]
                            ,[AccommodationStatusDescription]
                             FROM [NHSE_Reference].[dbo].[tbl_Ref_DataDic_ECDS_Accommodation_Status]")
data_accom <- data_accom %>% mutate(AccommodationStatusCode=as.character(AccommodationStatusCode))
care_accom <- c('394923006','160734000')
tb_accom <- data_accom %>% mutate(care = ifelse(AccommodationStatusCode %in% care_accom,1,0)) %>% arrange(desc(care))


### Load acuity

data_acu <- dbGetQuery(con,
                         "SELECT [AcuityCode]
                         ,[AcuityDescription]
                             FROM [NHSE_Reference].[dbo].[tbl_Ref_DataDic_ECDS_Acuity]")
data_acu <- data_acu %>% mutate(AcuityCode=as.character(AcuityCode))

```

```{r ECDS_data, include=FALSE}

data_EC_pl <- dbGetQuery(con,
                         "SELECT * FROM
                         [NHSE_Sandbox_EmpowerThePerson].[dbo].[MFonseca_EC_careproxy_2122v2_agg]")

```



```{r ECDS_data_preprocess, include=FALSE}


enrich <- function(data_EC_pl){
  
  data_EC_pl <- data_EC_pl %>% mutate(EC_Chief_Complaint_SNOMED_CT = as.character(EC_Chief_Complaint_SNOMED_CT),
                                      EC_Acuity_SNOMED_CT = as.character(EC_Acuity_SNOMED_CT))
  
  data_EC_pl_d <- data_EC_pl %>% left_join(data_complaint,by=c("EC_Chief_Complaint_SNOMED_CT"="ChiefComplaintCode"))
  
  data_EC_pl_d <- data_EC_pl_d %>% left_join(data_grouping,by=c("EC_Chief_Complaint_SNOMED_CT"="ChiefComplaintCode"))
  
  data_EC_pl_d <- data_EC_pl_d %>%
    left_join(data_acu, by=c("EC_Acuity_SNOMED_CT"="AcuityCode"))
  
}


data_EC_pl <- enrich(data_EC_pl)


# Create age groups, simplify gender, remove under 18 year olds

data_EC_pl_clean <- data_EC_pl %>%
  mutate(
    Der_Age_At_CDS_Activity_Date = as.integer(Der_Age_At_CDS_Activity_Date),
    age_group = factor(cut(Der_Age_At_CDS_Activity_Date,breaks=c(0,17,44,54,64,74,84,Inf))),
    age_65 = factor(cut(Der_Age_At_CDS_Activity_Date,breaks=c(0,17,64,Inf))),
    Sex = factor(case_when(Sex=="1" ~ "Male", Sex=="2" ~ "Female", TRUE ~ "Other/Unknown")),
    care_flag = ifelse(care_flag==1,"Care","Other")) %>%
  filter(Der_Age_At_CDS_Activity_Date>=18)

```



## Approach

In this rapid exploratory analysis (***not QA'd***), we have looked at the ECDS (emergency care dataset) activity in financial year 21/22.

In terms of identifying those in receipt of social care - full cohort identification may be prohibitive without external (likely local authority held patient-level information).
Identification of those in care homes is more feasible. A request has put in for access to the restricted master patient index (MPI) dataset, which includes a flag identifying care home residents based on NHSAI postcode information. This is not yet available though.

In the meantime, a proxy approach has been used to identify those in care homes (and to some extent other social care users) based on existing variables in ECDS.

The 'care_flag' was attributed if **either** one of the following was registered in the event:

* A relevant 'Discharge Destination' code, **or**
* A relevant 'Attendance Source' code, **or**
* A relevant Accommodation Status code' (in theory identifies those in care homes, but data quality/completion issues, hence need for MPI ideally)

The chosen relevant codes (judgemental) are indicated below, with a '1'.


```{r care_flag_ref}

kable(tb_source) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

kable(tb_discdest) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

kable(tb_accom) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))


```

## Results

When referring to care home resident in any of the results above, this is with respect to those identified with the proxy 'care_flag'.
Results cover all emergency care attendances in 21/22 for adults (18+).


### Pivot table on activity, care flag, age (65 or over), chief complaint

Some of the data, in aggregate form, is available below to inspect.
Aggregation is done based on chief complaint, care flag, age (65 and over or not). Cases where activity is 5 or under are suppressed and substituted with a value of 2.5.

```{r pivot_prep, fig.height=20, fig.width=20, include=FALSE}

data_pivot <- data_EC_pl_clean %>%
  group_by(care_flag,ChiefComplaintGrouping,ChiefComplaintDescription,age_65,Sex) %>%
  summarise(attendances=sum(n,na.rm=TRUE)) %>% ungroup() %>%
  mutate(attendances=ifelse(attendances<=5,2.5,attendances))
library(DT)

```


```{r pivot_1, fig.width=20,fig.height=20}


datatable(data_pivot,
          filter = 'top', 
          extensions = 'Buttons',
          options = list(dom = 'Blfrtip',
          buttons = c('copy', 'csv', 'pdf', 'print'),
          lengthMenu = list(c(10,25,50,-1),
          c(10,25,50,"All"))))

```



### Chief complaint grouping

Below the chief complaint grouping for each attendance is shown, disaggregated by the care_flag.

```{r plot_chief_prep, fig.height=20, fig.width=20, include=FALSE}
aux_ccg <- data_EC_pl_clean %>% group_by(care_flag,ChiefComplaintGrouping) %>% summarise(n0=sum(n)) %>% arrange(care_flag,desc(n0))
aux_ccg$ChiefComplaintGrouping <- factor(aux_ccg$ChiefComplaintGrouping,levels=unique(aux_ccg$ChiefComplaintGrouping))


data_EC_pl_clean$ChiefComplaintGrouping <- factor(data_EC_pl_clean$ChiefComplaintGrouping,unique(aux_ccg$ChiefComplaintGrouping))

```


```{r plot_chief, fig.width=20,fig.height=20}

ggplot(data = aux_ccg) +
  geom_bar(aes(x=ChiefComplaintGrouping,y=n0),stat="identity")+
  facet_wrap(~care_flag,ncol=1,scales="free")+
  labs(y="Count")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        text=element_text(size=22))


```

### Chief complaint description

Below the top 15 chief complaints for those with a care flag are shown (among ca. 142 codes).

```{r prep_ccd, fig.height=20, fig.width=20, include=FALSE}
aux <- data_EC_pl_clean %>% group_by(care_flag,ChiefComplaintDescription) %>% summarise(n0=sum(n)) %>% arrange(care_flag,desc(n0))
aux$ChiefComplaintDescription <- factor(aux$ChiefComplaintDescription,levels=unique(aux$ChiefComplaintDescription))

top15care <- aux %>% filter(!is.na(ChiefComplaintDescription)) %>% .$ChiefComplaintDescription %>% as.character() %>% .[1:15]


aux_sd <- data_EC_pl_clean %>%
  group_by(care_flag,ChiefComplaintDescription,age_group,Sex) %>%
  summarise(n0=sum(n)) %>%
  ungroup() %>%
  group_by(care_flag) %>%
  mutate(perc = n0/sum(n0)) %>% ungroup()
aux_sd$ChiefComplaintDescription <- factor(aux_sd$ChiefComplaintDescription,levels=unique(aux$ChiefComplaintDescription))
```

```{r plot_complaint, fig.width=20,fig.height=20}


ggplot(data = aux %>% filter(ChiefComplaintDescription %in% top15care,care_flag=="Care")) +
  geom_bar(aes(x=ChiefComplaintDescription,y=n0),stat="sum")+
  facet_wrap(~care_flag,ncol=1,scales="free")+
  labs(y="Count",title="Top chief complaints for those with care home flag")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        text=element_text(size=22))+
  coord_flip()


```

#### Chief complaint diagnosis - comparison to remaining cohort

Below the top 15 chief complaints for those with a care flag are shown, as relative frequency among the complaints. The relative frequency of these diagnoses for non-care cohort is also shown.

```{r plot_complaint_rel, fig.width=20,fig.height=20}


ggplot(data = aux_sd %>% filter(ChiefComplaintDescription %in% top15care),
       aes(x=ChiefComplaintDescription,y=perc*100,fill=care_flag)) +
  stat_summary(geom = "bar",fun="sum",position = position_dodge(0.95))+
  #facet_wrap(~care_flag,ncol=1,scales="free")+
  labs(y="% of all attendances for this cohort",title="Top Chief Complaints for those with social care flag, as % of all attendances for that cohort")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        text=element_text(size=22))+
  coord_flip()


```


#### Chief complaint diagnosis (care cohort) - breakdown by age groups

The breakdown by age is made explicit below.

```{r plot_complaint_rel_age, fig.width=20,fig.height=20}


ggplot(data = aux_sd %>% filter(ChiefComplaintDescription %in% top15care,care_flag=="Care"),
       aes(x=ChiefComplaintDescription,y=n0,fill=age_group)) +
  stat_summary(geom = "bar",fun="sum",position="stack")+
  #facet_wrap(~care_flag,ncol=1,scales="free")+
  labs(y="no attendances for this cohort",title="Top Chief Complaints for those with social care flag - age breakdown")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        text=element_text(size=22))+
  coord_flip()+
  scale_fill_brewer(palette = "BrBG")
  #theme_economist()

```

#### Chief complaint diagnosis (care cohort) - breakdown by gender

Breakdown by gender is shown.

```{r plot_complaint_rel_sex, fig.width=20,fig.height=20}


ggplot(data = aux_sd %>% filter(ChiefComplaintDescription %in% top15care,care_flag=="Care"),
       aes(x=ChiefComplaintDescription,y=n0,fill=Sex)) +
  stat_summary(geom = "bar",fun="sum",position="stack")+
  #facet_wrap(~care_flag,ncol=1,scales="free")+
  labs(y="no attendances for this cohort",title="Top Chief Complaints for those with social care flag - sex breakdown")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        text=element_text(size=22))+
  coord_flip()

```







### Activity by emergency care department type

The breakdown by emergency care department type is shown below. NA reflects where data is unavailable.


```{r plot_ECD, fig.width=15,fig.height=15}

ggplot(data = data_EC_pl_clean,
       aes(x=Pat_Der_Dimention_1,y=n,fill=care_flag)) +
  stat_summary(geom = "bar",fun="sum",position="stack")+
  facet_wrap(~care_flag,scales="free_x")+
  labs(y="no attendances for this cohort",title="Emergency care department breakdown")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        text=element_text(size=22))+coord_flip()


ggplot(data = data_EC_pl_clean %>% filter(Der_Age_At_CDS_Activity_Date>=65),
       aes(x=Pat_Der_Dimention_1,y=n,fill=care_flag)) +
  stat_summary(geom = "bar",fun="sum",position="stack")+
  facet_wrap(~care_flag,scales="free_x")+
  labs(y="no attendances for this cohort",title="Emergency care department breakdown, 65+ only",x="")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        text=element_text(size=22))+coord_flip()

```


### Activity by acuity

The breakdown by acuity is shown below. NA reflects where data is unavailable.


```{r plot_acuity, fig.width=15,fig.height=15}

ggplot(data = data_EC_pl_clean,
       aes(x=AcuityDescription,y=n,fill=care_flag)) +
  stat_summary(geom = "bar",fun="sum",position="stack")+
  facet_wrap(~care_flag,scales="free_x")+
  labs(y="no attendances for this cohort",title="Acuity breakdown")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        text=element_text(size=22))+coord_flip()


ggplot(data = data_EC_pl_clean %>% filter(Der_Age_At_CDS_Activity_Date>=65),
       aes(x=AcuityDescription,y=n,fill=care_flag)) +
  stat_summary(geom = "bar",fun="sum",position="stack")+
  facet_wrap(~care_flag,scales="free_x")+
  labs(y="no attendances for this cohort",title="Acuity breakdown, 65+",x="")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        text=element_text(size=22))+coord_flip()

```

### Outcome

The breakdown by outcome from attendance is shown below. NA reflects where data is unavailable.


```{r plot_outcome, fig.width=15,fig.height=15}

ggplot(data = data_EC_pl_clean,
       aes(x=Pat_Der_Dimention_6,y=n,fill=care_flag)) +
  stat_summary(geom = "bar",fun="sum",position="stack")+
  facet_wrap(~care_flag,scales="free_x")+
  labs(y="no attendances for this cohort",title="Outcome breakdown",x="Outcome")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        text=element_text(size=22))+coord_flip()


ggplot(data = data_EC_pl_clean %>% filter(Der_Age_At_CDS_Activity_Date>=65),
       aes(x=Pat_Der_Dimention_6,y=n,fill=care_flag)) +
  stat_summary(geom = "bar",fun="sum",position="stack")+
  facet_wrap(~care_flag,scales="free_x")+
  labs(y="no attendances for this cohort",title="Outcome breakdown, 65+",x="Outcome")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        text=element_text(size=22))+coord_flip()

```


